# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Non Personalized Recommendation Engine

Author: Adolfo de Rueda Panadero

========================================================

# Initialization

Import libraries

```{r}

library(reshape)
library(tidyr)

```

Load the critics dataset.

```{r}
critics = read.csv("critics.csv", sep = ",")
```

# 1) Top 5 Movies, by rating mean
Calculate mean rating for each movie, ordered with the highest rating listed first, and plot the top 5.

**Expected result**:
3.600000, 'The Shawshank Redemption'
3.266667, 'Star Wars IV - A New Hope'
3.222222, 'Blade Runner'
3.166667, 'Groundhog Day'
3.062500, 'The Silence of the Lambs'

```{r fig.width=7, fig.height=6}

#get the mean of each column into a matrix
means_col = matrix(colMeans(critics[-1], na.rm = TRUE))
#get the names of each column
film_col = matrix(colnames(critics[-1]))
#bind them into the same matrix
ratings = cbind(means_col,film_col)
#print top 5
head(ratings[order(ratings[,1], decreasing = TRUE),], n = 5)

```

# 2) Top 5 Movies, by rating distribution
Calculate the percentage of ratings for each movie that are 4 stars or higher. Order with the highest percentage first, and plot the top 5 movies with more high ratings distribution.

**Expected result**:
0.7000000, 'The Shawshank Redemption'
0.5333333, 'Star Wars IV - A New Hope'
0.5000000, 'Gladiator'
0.4444444, 'Blade Runner'
0.4375000, 'The Silence of the Lambs'

```{r fig.width=7, fig.height=6}
#create two matrixes empty for the next for loop
percentages = matrix(NA, ncol = 2, nrow = 20)
value = matrix(NA, ncol = 2)

#for loop to bind into a matrix two columns: first one sum the number of ratings >= 4 and divide them by the total and the name of the movie where you are carrying this operation
for(i in 2:ncol(critics)){
  x = i-1
  value = cbind(c(sum(critics[[i]]>=4, na.rm = TRUE)/sum(critics[[i]]>0, na.rm = TRUE)),c( colnames(critics[i]))) 
  percentages[x,] = value
}
#print top 5
head(percentages[order(percentages[,1], decreasing = TRUE),], n = 5)


```

# 3) Top 5 Movies, by quantity of ratings
Count the number of ratings for each movie, order the one with most number of ratings first, submit the top 5.

**Expected result**:
17, 'Toy Story'
16, 'The Silence of the Lambs'
15, 'Star Wars IV - A New Hope'
14, 'Star Wars VI - Return of the Jedi'
13, 'Independence Day'

```{r fig.width=7, fig.height=6}
#create two matrixes empty for the next for loop
nratings = matrix(NA, ncol = 2, nrow = 20)
value = matrix(NA, ncol = 2)

#for loop to bind into a matrix two columns: first one sum the total number of ratings and the name of the movie where you are carrying this operation
for(i in 2:ncol(critics)){
  x = i-1
  value = cbind(sum(critics[[i]]>=0, na.rm = TRUE),c(colnames(critics[i])))
  nratings[x,] = value
}

#In this case we need to make it data frame to modify the number of ratings column to make it number type data and the plot the top 5
nratings = data.frame(nratings)
nratings$X1 = as.numeric(as.character(nratings$X1))
str(nratings)
head(nratings[order(nratings[,1], decreasing = TRUE),], n = 5)

```

# 4) People who watched Star Wars IV also watched ...
Calculate movies that most often occur with other movie. For each movie, calculate the percentage of the other movie raters who also rated that movie. Order with the highest percentage first, and submit the top 5. Select "Star.Wars.IV...A.New.Hope" movie by defaut. Use the (x+y)/x method.

Top Occurences:
0.9333333, 'Toy Story'
0.8666667, 'Star Wars: Episode VI - Return of the Jedi'
0.8000000, 'The Silence of the Lambs'
0.7333333, 'Independence Day'
0.6666667, 'Total Recall'

```{r fig.width=7, fig.height=6}
#create two matrixes empty for the next for loop
preferences = matrix(NA, ncol = 2, nrow = 20)
ratio = matrix(NA, ncol = 2)

#for loop to bind into a matrix two columns: first one sum the number of ratings that "Star Wars IV: A New Hope" and the other movies coincided and the name of the movie where you are carrying this operation
for(i in 3:nrow(critics)){
  x = i-1
  ratio = cbind(sum((is.na(critics[[i]])==FALSE) & (is.na(critics[[2]])==FALSE), na.rm = TRUE)/sum(((is.na(critics[[2]])==FALSE))), c(colnames(critics[i])))
  preferences[x,] = ratio
}

#print top 5
head(preferences[order(preferences[,1], decreasing = TRUE),], n = 5)

```


# 5) People who liked Babe also liked ...
Calculate the movies better rated of people who liked a movie. Select the people who liked the movie "Babe" (4 or 5 stars) and provide the top 5 movies they liked most.

**Expected result**:
5.000, 'Pulp.Fiction'
5.000, 'Groundhog.Day'
4.500, 'The.Shawshank.Redemption'
4.333, 'Toy.Story'
4.000, 'Blade.Runner'

```{r fig.width=7, fig.height=6}

#create two matrixes empty for the next for loop and a new dataset with only the rows where "Babe" ratings are 4 or higher
liked_most = matrix(NA, ncol = 2, nrow = 20)
ratio = matrix(NA, ncol = 2)
new_data = critics[critics[,21]>=4,]
new_data = new_data[complete.cases(new_data[,21]),]

# in this for loop we create a two column matrix: first column with the average rate for each movie and the second with the name of the movie
for(i in 2:ncol(critics)-1){
  x = i-1
  ratio = cbind(mean(new_data[,i], na.rm = TRUE), colnames(new_data[i]))
  liked_most[x,] = ratio
}

# print the top 5
head(liked_most[order(liked_most[,1], decreasing = TRUE),], n = 5)



```

# 6) MovieLens
Explore a real data set and provide non-personalized ratings. You can use the MovieLens database. You can find MovieLens' dataset here: http://files.grouplens.org/datasets/movielens/ml-10m-README.html

## Small

```{r}

# Loading the data

movies_100k = read.csv("movies.csv", sep = ",")

ratings_100k = read.csv("ratings.csv", sep = ",")

# Putting both databases together

data_100k = merge(ratings_100k, movies_100k, by.x = 'movieId')

#Erasing columns we dont need due to merge
data_100k$movieId = NULL
data_100k$genres = NULL
data_100k$timestamp = NULL

# We use spread function to give our data the same structure as in past exercises
data_100k = data_100k[-86634,]#repeated rating --> user rates the movie twice
data_100k = spread(data = data_100k, key = title, value = rating)
#head(data_100k) --> It takes a little bit of time


```


# Exercise 1

Calculate mean rating for each movie, ordered with the highest rating listed first, and plot the top 5.

```{r}
# Same as Exercise 1

means_col_100 = matrix(colMeans(data_100k[-1], na.rm = TRUE))
film_col_100 = matrix(colnames(data_100k[-1]))
ratings_100 = cbind(means_col_100,film_col_100)
head(ratings_100[order(ratings_100[,1], decreasing = TRUE),], n = 5)


```

# Exercise 2

Calculate the percentage of ratings for each movie that are 4 stars or higher. Order with the highest percentage first, and plot the top 5 movies with more high ratings distribution.

```{r}
# Same as exercise 2

percentages = matrix(NA, ncol = 2, nrow = ncol(data_100k))
value = matrix(NA, ncol = 2)

for(i in 2:ncol(data_100k)){
  x = i-1
  value = cbind(c(sum(data_100k[,i]>=4, na.rm = TRUE)/sum(data_100k[,i]>0, na.rm = TRUE)),c( colnames(data_100k[i]))) 
  percentages[x,] = value
  
}

head(percentages[order(percentages[,1], decreasing = TRUE),], n = 5)


```

# Exercise 3

Count the number of ratings for each movie, order the one with most number of ratings first, submit the top 5.
```{r}
# Same as exercise 3

nratings = matrix(NA, ncol = 2, nrow = ncol(data_100k))
value = matrix(NA, ncol = 2)

for(i in 2:ncol(data_100k)){
  x = i-1
  value = cbind(sum(data_100k[[i]]>=0, na.rm = TRUE),c(colnames(data_100k[i])))
  nratings[x,] = value
}

nratings = data.frame(nratings)
nratings$X1 = as.numeric(as.character(nratings$X1))
nratings = as.matrix(nratings)

head(nratings[order(nratings[,1], decreasing = TRUE),], n = 5)


```

# Exercise 4
Calculate movies that most often occur with other movie. For each movie, calculate the percentage of the other movie raters who also rated that movie. Order with the highest percentage first, and submit the top 5. Use the (x+y)/x method.

We will select Bottle Rocket as default. 

```{r}

#Same as in exercise 4

preferences = matrix(NA, ncol = 2, nrow = nrow(data_100k))
ratio = matrix(NA, ncol = 2)

for(i in 3:nrow(data_100k)){
  x = i-1
  ratio = cbind(sum((is.na(data_100k[[i]])==FALSE) & (is.na(data_100k[[2]])==FALSE), na.rm = TRUE)/sum(((is.na(data_100k[[2]])==FALSE))), c(colnames(data_100k[i])))
  preferences[x,] = ratio
}

head(preferences[order(preferences[,1], decreasing = TRUE),], n = 5)

```

# Exercise 5
Calculate the movies better rated of people who liked a movie. Select the people who liked the movie "Forrest Gump (1994)" (4 or 5 stars for the last movie on the dataset) and provide the top 5 movies they liked most.

```{r}

#Same as in exercise 5

liked_most = matrix(NA, ncol = 2, nrow = ncol(data_100k))
ratio = matrix(NA, ncol = 2)
new_data = data_100k[data_100k$`Forrest Gump (1994)`>=4,]
new_data = new_data[complete.cases(new_data$`Forrest Gump (1994)`),]

for(i in 2:ncol(data_100k)){
  x = i-1
  ratio = cbind(mean(new_data[,i], na.rm = TRUE), colnames(new_data[i]))
  liked_most[x,] = ratio
}

liked_most = liked_most[liked_most[,1]!="NaN",]
head(liked_most[order(liked_most[,1], decreasing = TRUE),], n = 5)

```

