# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: Adolfo de Rueda Panadero

========================================================

# Initialization

Import libraries

```{r}


```

Load the critics dataset.

```{r}
critics = read.csv("C:/Users/adolf/OneDrive/Documents/IE Business School/2nd Semester/Recommendation Engines/IE-Lab1/Data/critics.csv", sep = ",")
```

# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Victoria and Nuria

**Expected correlation between Victoria and Nuria**: 0.3651484

```{r fig.width=7, fig.height=6}

```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Victoria and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?

# 3) Top Recommendations 
Get the top 5 movies recommended for Victoria. by using a weighted average of every other critic's rankings.

**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'

```{r fig.width=7, fig.height=6}

```

# 4) Top similar critics
Return the top 5 critics that are most similar to Victoria. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}
```